Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-11 paper.tex
the CWA -> CWA

The contributions of this paper are (a) formal re-engineering of the CWA
(b) providing an additional security and privacy analysis with interactive theorem proving
certification of a novel view on the system architecture including actors, locations, and policies,
(c) a formal definition of a security refinement process that allows to improve a system
based on attacks found by the attack tree analysis and (d) an application of the refinement to
improve security of the CWA specification.

In this paper, we first provide some background in Section \ref{sec:background}:
we give a brief overview of related works
and the protocol of CWA (Section \ref{sec:history}).
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on the CWA (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

\section{Background and related work}
\label{sec:background}
\subsection{DP-3T and PEPP-PT}
\label{sec:history}
We are mainly concerned with the architecture and protocols proposed by the
DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project \cite{dp3t:github}.
The main reason to focus on this particular family of protocols is the
\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google \cite{enf:proj}, that adopts
%the main
core principles of the DP-3T proposal. This API is not only used in the German
CWA but has the potential of being widely adopted in future
app developments that might emerge due to the reach of players like Apple and Google.

@@

\subsubsection{Instantiation of Framework}
The formal model of CWA uses the Isabelle Infrastructure framework instantiating it 
by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
to physical locations. The Ephemeral Ids, their sending and change is added to Infrastructures
by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules
for the actions move, get, and put. The details of the newly adapted Infrastructure are
presented in Section \ref{sec:model}.
Technically, an Isabelle theory file \texttt{Infrastructure.thy} builds on top of the theories for Kripke 
structures and CTL (\texttt{MC.thy}), attack trees (\texttt{AT.thy}), and security refinement 
(\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
for the CWA, express relevant and interesting properties, and conduct
interactive proofs (with the full support of the powerful and highly automated proof support
of Isabelle).
%The CWA theory itself is an adaptation of the Infrastructure theory of the 
%Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
%In the remainder of this paper, we introduce the model that we conceived for . 
All Isabelle sources are available online \cite{kam:20gitsc}.

@@

In the CWA, the initial values for the \texttt{igraph} components use
two locations \texttt{pub} and \texttt{shop} to define the following constants (we omit the
data map component \texttt{ex\_locs}).
\begin{ttbox}
ex_loc_ass \ttequiv (\ttlam x. if x = pub then \{''Alice'', ''Bob'', ''Eve''\}  
                 else (if x = shop then \{''Charly'', ''David''\} 
                 else \{\}))
ex_creds \ttequiv (\ttlam x. if x = ''Alice'' then (\{\}, \{\}, Efid 1) else 
              (if x = ''Bob'' then  (\{\},\{\}, Efid 2) else 
              (if x = ''Charly'' then (\{\},\{\}, Efid 3) else
              (if x = ''David'' then (\{\},\{\}, Efid 4) else
              (if x = ''Eve'' then (\{\},\{\}, Efid 5)
               else (\{\},\{\},Efid 0))))))
ex_efids \ttequiv (\ttlam x. if x = pub then \{Efid 1, Efid 2, Efid 5\}
                  else (if x = shop then \{Efid 3, Efid 4\} else \{\}))
ex_knos \ttequiv (\ttlam x. (if x = Actor ''Eve'' then (\ttlam l. \{\} else (\ttlam l. \{\})))
\end{ttbox}  
These components are wrapped up into the following \texttt{igraph}.
\begin{ttbox}
ex_graph \ttequiv
      Lgraph \{(pub, shop)\} ex_loc_ass ex_creds ex_locs ex_efids ex_knos
\end{ttbox}  
Infrastructures are given by the following datatype that 
contains an infrastructure graph of type \texttt{igraph} 
and a policy given by a function that assigns local policies over a graph to
all locations of the graph. 
\begin{ttbox}
{\bf{datatype}} infrastructure = Infrastructure igraph 
                                         [igraph, location] \ttfun policy set
\end{ttbox}
There are projection functions \texttt{graphI} and \texttt{delta} when applied
to an infrastructure return the graph and the local policies, respectively.
The function \texttt{local\_policies} gives the policy for each location \texttt{x}
over an infrastructure graph \texttt{G} as a pair: the first element of this pair is 
a function specifying the actors \texttt{y} that are entitled to perform the actions 
specified in the set which is the second element of that pair.
The local policies definition for the CWA, simply permits all actions to
all actors in both locations.
\begin{ttbox}
  local_policies G \ttequiv
         (\ttlam x. if x = pub then  \{(\ttlam y. True, \{get,move,put\})\}
          else (if x = shop then \{(\ttlam y. True, \{get,move,put\})\} else \{\}))
\end{ttbox}  
For the CWA, the initial infrastructure contains the graph \texttt{ex\_graph}
with its two locations pub and shop and is then wrapped up with the local policies
into the infrastructure \texttt{Corona\_scenario} 
that represents the ``initial'' state for the Kripke structure.
\begin{ttbox}
 Corona_scenario \ttequiv Infrastructure  ex_graph local_policies
\end{ttbox}

@@

In addition, we also have a simple implicit version of {\it action refinement}. In an
action refinement, traces may be refined by combining consecutive system events
into atomic events thereby reducing traces.
We can observe this kind of refinement in the second refinement step
%This will be detailed in the second refinement step
of the CWA considered next.

>>>>

The contributions of this paper are (a) formal re-engineering of CWA
(b) providing an additional security and privacy analysis with interactive theorem proving
certification of a novel view on the system architecture including actors, locations, and policies,
(c) a formal definition of a security refinement process that allows to improve a system
based on attacks found by the attack tree analysis and (d) an application of the refinement to
improve security of the CWA specification.

In this paper, we first provide some background in Section \ref{sec:background}:
we give a brief overview of related works
and the protocol of CWA (Section \ref{sec:history}).
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on CWA (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

\section{Background and related work}
\label{sec:background}
\subsection{DP-3T and PEPP-PT}
\label{sec:history}
We are mainly concerned with the architecture and protocols proposed by the
DP-3T (\textit{Decentralized Privacy-Preserving Proximity Tracing}) project \cite{dp3t:github}.
The main reason to focus on this particular family of protocols is the
\textit{Exposure Notification Framework} (ENF), jointly published by Apple and Google \cite{enf:proj}, that adopts
%the main
core principles of the DP-3T proposal. This API is not only used in 
CWA but has the potential of being widely adopted in future
app developments that might emerge due to the reach of players like Apple and Google.

@@

\subsubsection{Instantiation of Framework}
The formal model of CWA uses the Isabelle Infrastructure framework instantiating it 
by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
to physical locations. The Ephemeral Ids, their sending and change is added to Infrastructures
by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules
for the actions move, get, and put. The details of the newly adapted Infrastructure are
presented in Section \ref{sec:model}.
Technically, an Isabelle theory file \texttt{Infrastructure.thy} builds on top of the theories for Kripke 
structures and CTL (\texttt{MC.thy}), attack trees (\texttt{AT.thy}), and security refinement 
(\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
for CWA, express relevant and interesting properties, and conduct
interactive proofs (with the full support of the powerful and highly automated proof support
of Isabelle).
%The CWA theory itself is an adaptation of the Infrastructure theory of the 
%Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
%In the remainder of this paper, we introduce the model that we conceived for . 
All Isabelle sources are available online \cite{kam:20gitsc}.

@@

In CWA, the initial values for the \texttt{igraph} components use
two locations \texttt{pub} and \texttt{shop} to define the following constants (we omit the
data map component \texttt{ex\_locs}).
\begin{ttbox}
ex_loc_ass \ttequiv (\ttlam x. if x = pub then \{''Alice'', ''Bob'', ''Eve''\}  
                 else (if x = shop then \{''Charly'', ''David''\} 
                 else \{\}))
ex_creds \ttequiv (\ttlam x. if x = ''Alice'' then (\{\}, \{\}, Efid 1) else 
              (if x = ''Bob'' then  (\{\},\{\}, Efid 2) else 
              (if x = ''Charly'' then (\{\},\{\}, Efid 3) else
              (if x = ''David'' then (\{\},\{\}, Efid 4) else
              (if x = ''Eve'' then (\{\},\{\}, Efid 5)
               else (\{\},\{\},Efid 0))))))
ex_efids \ttequiv (\ttlam x. if x = pub then \{Efid 1, Efid 2, Efid 5\}
                  else (if x = shop then \{Efid 3, Efid 4\} else \{\}))
ex_knos \ttequiv (\ttlam x. (if x = Actor ''Eve'' then (\ttlam l. \{\} else (\ttlam l. \{\})))
\end{ttbox}  
These components are wrapped up into the following \texttt{igraph}.
\begin{ttbox}
ex_graph \ttequiv
      Lgraph \{(pub, shop)\} ex_loc_ass ex_creds ex_locs ex_efids ex_knos
\end{ttbox}  
Infrastructures are given by the following datatype that 
contains an infrastructure graph of type \texttt{igraph} 
and a policy given by a function that assigns local policies over a graph to
all locations of the graph. 
\begin{ttbox}
{\bf{datatype}} infrastructure = Infrastructure igraph 
                                         [igraph, location] \ttfun policy set
\end{ttbox}
There are projection functions \texttt{graphI} and \texttt{delta} when applied
to an infrastructure return the graph and the local policies, respectively.
The function \texttt{local\_policies} gives the policy for each location \texttt{x}
over an infrastructure graph \texttt{G} as a pair: the first element of this pair is 
a function specifying the actors \texttt{y} that are entitled to perform the actions 
specified in the set which is the second element of that pair.
The local policies definition for CWA, simply permits all actions to
all actors in both locations.
\begin{ttbox}
  local_policies G \ttequiv
         (\ttlam x. if x = pub then  \{(\ttlam y. True, \{get,move,put\})\}
          else (if x = shop then \{(\ttlam y. True, \{get,move,put\})\} else \{\}))
\end{ttbox}  
For CWA, the initial infrastructure contains the graph \texttt{ex\_graph}
with its two locations pub and shop and is then wrapped up with the local policies
into the infrastructure \texttt{Corona\_scenario} 
that represents the ``initial'' state for the Kripke structure.
\begin{ttbox}
 Corona_scenario \ttequiv Infrastructure  ex_graph local_policies
\end{ttbox}

@@

In addition, we also have a simple implicit version of {\it action refinement}. In an
action refinement, traces may be refined by combining consecutive system events
into atomic events thereby reducing traces.
We can observe this kind of refinement in the second refinement step
%This will be detailed in the second refinement step
of CWA considered next.
